Which ``Bad Smells'' can be Ignored?

### 2016-03-11 Manuscript.tex
Updates from ShareLaTeX

\documentclass{sig-alternate}
\usepackage{stfloats}
\usepackage{tikz}
  \def\firstcircle{(90:1.75cm) circle (2.5cm)}
  \def\secondcircle{(210:1.75cm) circle (2.5cm)}
  \def\thirdcircle{(330:1.75cm) circle (2.5cm)} 
\usepackage{comment}
\usepackage{cite}
\usepackage{framed,graphicx,xcolor}

\usepackage{rotating}
%\usepackage{stfloats}

@@

This paper proposes   XTREE, a tool to  prioritize bad smells. Given an historical
log of defects
seen previously in the code, XTREE
  deprioritizes   bad smells  that have little to no support
  in that log.

  This paper compares the bad smells found by XTREE to
  those found by previous work by Shatnawi, Alves, and Borges. 
  If code modules are  changed to avoid the bad
  smells found by XTREE,  that changed system
  has significantly fewer expected defects.
  Also, XTREE   endorses
  changes to far fewer code metrics.
  Hence, it is 
  a better tool for ruling out irrelvant bad smells.
  For example, the XTREE results show that many of the bad smells learned 
  by Shatnawi, Alves, and Borges can be ignored.

@@

On the other hand, other papers cast doubts on the value of use bad smalls
as triggers for code change.
As we show in this paper,  
 tools and text books and developers can disagree on what bad smells
are important enough to hunt down and eliminate. Our findings
are consistent with several other papers.
Mantyla et al.~\cite{Mantyla2004} found that within an organization,
 developers had different perceptions on the importance of different
 bad smells.    Yamashita  and Moonen~\cite{Yamashita2013}
 surveyed 85 professional developers to conclude that it remains
 open to debate if code smells are a meaningful conceptualization  of code quality issues (from a developer's perspective). 
 Sj\"oberg et al.~\cite{Sjoberg2013} studied six developers 
 performing maintenance tasks that took three to four weeks to complete. 
 After modifying nearly 300 Java files, it was observed
 that none of the dozen investigated code smells explored in 
 that study were mostly irrelevant to to that maintenance work.

@@

It turns out that it is hard to learn ``good'' bad smell detectors.
As discussed on the next page, programmer
   cognitive biases result in different
developers declaring that
bad smell X or Y is most important.

Also, a bad smell  detector
is a combination of some code metric (e.g. lines of code in a method)
and a threshold measurement (e.g. {\em loc} $>$ 100). 
Shantnawi~\cite{Shatnawi10} warn that  there are most  studies do not offer
{\em both} a metrics and a threshold for their proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
discuss what OO design metrics might defect bad smells (e.g. a excessive
depth of inheritance tree) but they never comment on what thresholds
means that some section of code starts to smell bad. After
an extensive literature review on OO bad smells, Shantnawi remarks that there is a dearth
of  studies  conducted to formulate bad smell guidelines,
represented as metrics plus threshold values, that interpret the complexity of the software design using OO code metrics.

Further, it is hard to judge the  effects of removing bad smells
for three reasons.
Firstly, the bad smell learners currently found in the literature
report bad smells for each code metric {\em in isolation}
Refactored code cannot be assessed just by a rerun of the the test
suite since
code refactorings is meant to
preserve the semantics of the code (i.e. the same tests
should pass before and after the refactoring). 
It is also problematic to assessing refactorings by making the
refactorings,  then seeing what happens to that code for the rest
of the project. In an test-driven agile project, as
SCRUM meetings change what stories are implemented next,
the operational profile that lead to defects in the old
code may not repeat.

To address the first problem of programmer cognitive biases,
this paper proposes using the   XTREE contrast set learner to
find bad smells. Contrast set learners   recommend {\em what to change} in order to move from
     one classification to another (e.g. from defective=yes to
     defective=no).
     When applied to defect predictors from static code measures,
that delta is a change to static code measures;
i.e. a code refactoring. Given an historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have little to no support in that log.

Note that XTREE also addresses the second problem since it reports
code metrics and useful thresholds that

To resolve this third problem, this paper adopts
a technique used by Harman et al.

Why are there so few papers offer bad smell metrics and thresholds? For the few that
do exist, why do developers disagree on the relevance of those bad smells?

When developers cannot fully assess the value of bad smells,
it is possible to use automatic means.

This paper proposes the XTREE contrast set learner for recognizing
``bad'' bad smells. 
Contrast set learners   recommend {\em what to change} in order to move from
     one classification to another (e.g. from defective=yes to
     defective=no).
     When applied to defect predictors from static code measures,
that delta is a change to static code measures;
i.e. a code refactoring. Hence, constrast set learners
like XTREE can be used to  recognize ``bad''
bad smells:
\begin{itemize}
    \item Consider $N$ developers each proposing a software refactoring 
based on a set of bad smells $S$.
\item XTREE will (a)~learn contrast sets
from the historical record; then (b)~reject  any refactoring based on ``bad'' bad smells, i.e.
 on any member of $S_{\mathit{bad}} \in S$
that XTREE could {\em not} be found in the historical record.
\end{itemize}
Note that we    recommend  XTREE  is used to critize, but  not to create,
bad smells. 
XTREE uses frequency counts of code features seen in the historical record\footnote{From LL: Note that ``new'' bad smells may not have a historical record. A limitation to be acknowledged and a reason why developer knowledge is still relevant.}
to generate
predictions about the effects of changing code
attributes. Such predictions may be inaccurate due to the infamous correlation-vs-causation problem
where spurious correlations can suggest effects that are not
truly causal.

XXX

Nevertheless, XTREE is a still a useful tool to critique and prune
irrelevant bad smells since, 
as show below, many of the bad smell detectors offered
in the literature (and supported by readily available tools)
fall outside of  space of
effects found  using XTREE's analysis.
That is,   XTREE can find 
{\em neither} correlation {\em nor}
causal evidence that those bad smells
have any support   in the historical record.\footnote{From LL: You need to say what ``support from the historical record'' is early in the paper. Are you mapping code smells in source files to defects? Code smells to performance regressions? Whatever it is, give an example.}

XXXX

The same cannot be said for other methods that try to learn
bad smell thresholds. \tion{outs}
discusses {\em outlier} methods and {\em centroid delta} methods
for learning bad smell thresholds.
The {\em  outlier methods}  used by
Shatnawi, Alves and Herman et al.~\cite{Shatnawi10,Alves2010,hermans15}
propose thresholds on all code metrics with outlier values.
Similarly, the older {\em centroid delta} methods  used in our 
prior work~\cite{me12c} also returns thresholds on most of the code
metrics. That is, outlier methods and centroid deltas methods
seem to be saying that many bad smell thresholds are possible.

we proposed CD, short for
``Centroid Deltas'', as a way to find bad smell thresholds:
\begin{itemize}
    \item Clustered
data and generated centroids for each cluster;
\item Computed contrast sets (i.e. the bad smells)
as the delta between attributes in adjacent centroids $C_1 - C_2$
where examples in $C_1$ had more defects than $C_2$.
\end{itemize}
As shown below, Centroid Deltas
generates change
recommendations that are far less effective than XTREE.
Worse still Centroid Deltas reported bad smells
for far more attributes than XTREE. Hence, that prior work
is less useful than XTREE at judging the  bad smells proposed by developers.

XTREE   has certain limits. Firstly, it is an oracle that  might reject   refactorings proposed
by developers. XTREE does not perform those refactorings.
Secondly, XTREE assumes   Fowler's bad smells
should be assessed via their contribution to software defects
(evidence: see the section in italics of the quote
that starts this introduction). Some researchers 
 comment that   debating bad smells is  useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
debates about refactorings can   (a)~introducing new comers to the code base or (b)~share knowledge around a team
about interactions within the code. In that view, 
the concept of refactoring can be useful
even if it does not address issues of defects. 
That said, using
defect predictors,
it is possible to operationalize XTREE
while it might be harder to otherwise operationalize the assessment of bad smells.
Hence, 
this paper assumes that {\em better} bad smells and
bettering at finding code changes that  {\em reducing} the chances
that the new code will contain defects.

The contributions of this paper are
\begin{itemize}

    \item A new kind of contrast set learner called XTREE
    that   critique plans of what to change   in order
    to improve outcomes.
    \item An experimental rig for assessing the value of tools like
    XTREE that propose changes to software. That rig uses data mining
    to build an oracle for the proposed changes.
    \item Using that learner and that oracle,
    a demonstration that XTREE is better than our prior approach~\cite{me12c}
    as well as   state of the art methods~\cite{Shatnawi10,Alves2010,hermans15}
    at avoiding refactorings that are irrelevant to the task of reducing the risk of introducing defects.
    \end{itemize}
    Based on the above, we can offer
    a resolution between proponents of bad smells (e.g. Fowler, Shantnawi, Tufano~\cite{fowler99,Shatnawi10,Tufano2015}) and those who find them not useful or confusing   (e.g.Mantyla, Sj\"oberg,Yamahista et al.~\cite{Sjoberg2013,Mantyla2004,Yamashita2013}). We say
    that bad smells can guide code change, but only if those bad smells are
    first carefully assessed.

The rest of this paper discusses the cognitive
biases which make developers disagree on the relevance of those bad smells to
their particular projects.  
 That discussion is our motivation for developing the XTREE  tool
that   carefully assesses the merit of old bad smells to new projects.  XTREE is presented in Section 3 and evalauted in the rest of the paper.

@@

Two of the studies in \fig{smells} offers some comments on the relative importance
of the different bad smells the   Yamashita'13 study and the developer study:
    \begin{itemize}
        \item Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey. Those three were {\em duplicated code, large class}, 
        and {\em long method}.
        \item
 Note that this agreement also means that the
  Yamashita'13 study and the {\em developer survey}   concluded
        that most of the  listed code smells are {\em not} high priority issues
        requiring refactoring. 
 \end{itemize} 
In summary, just because a developer strongly believes something
about bad smells, does
not necessary mean that that belief is useful for the current project.
Human developers can be clever, but their thinking can also be distorted
by cognitive biases.
Hence, as shown in \fig{smells}, developers and text books and tools 
can disagree on which bad smells are important.
Hence it is important to have some automated tool
that can assess their beliefs in, for example,
bad smells.

\section{Learning Bad Smell Thresholds}
The SE literature offers  two ways to 
learning bad smell thresholds:
\begin{itemize}
    \item Approaches based on {\em outlier statistics}~\cite{erni96,bender99}
    as used by Shatnawi~\cite{Shatnawi10}, Alvenes et al.~\cite{Alves2010}
    and Hermans et al.~\cite{hermans15};
    \item  Approaches based on {\em cluster deltas} that we developed
    for   Centroid Deltas~\cite{me12c} and XTREE.
\end{itemize}
\subsection{  Outliers}

@@

\subsubsection{Enri \& Lewerentz}
One of the first methods was proposed Erni and Lewerentz~\cite{erni96}. 
Given classes described with the  code metrics of \fig{ck},
Enri and Lewerentz finds the   mean $\mu$ and the standard deviation $\sigma$
of each
code metrics. Their definition of problematic outlier was any code
metric with a measurement greater than $\mu+\sigma$.

Shatnami and Alves et al.~\cite{Shatnawi10,Alves2010} depreciate
using $\mu+\sigma$. since it does not   consider the fault-proneness of classes when the thresholds are computed. Also, there is a lack of empirical validation of this method.

\subsubsection{ Shatnami}
Shatnami~\cite{Shatnawi10}'s preferred alternative to $\mu+\sigma$
is to use the VARL method (Value of Acceptable Risk Level) proposed by Bender~\cite{bender99} in his epidemiology studies.  This approach uses two
thresholds which, following Shatnami's guidance, we set to
$p_0=p_1=0.05$. VARL encodes the defect count
for each class to 0 (for no defects known) to 1 (for defects known).
Univariate binary logistic regression is applied to learn three coefficients: $\alpha, \beta,p_0$ being
the  constant intercept ($\alpha$), the coefficient for maximizing log-likelihood function ($\beta$),
and a measure of how well this regression model predicts for the defects ($p_0$).
Any code metric with $p_0>0.05$ is then rejected as being a poor defect predictor. Thresholds are then learned from the surviving metrics $M_c$ using
the equation proposed by Bender:
\begin{equation}
 \mathit{bad\; smell\; if\;} M_c > \frac{1}{\beta }\left( {\log \left( {\frac{{{p_1}}}{{1 - {p_1}}}} \right) - \alpha } \right) 
\end{equation}
\subsubsection{ Alves et al.}
Alves et al.~\cite{Alves2010} propose another approach
to finding thresholds that  utilizes the underlying statistical distribution and scale of the metrics. 
Evey metric value is weighted according to the source lines of code (LOC) of the class. All the weighted metrics are then normalized; i.e. they are divided by the sum of all weights of the same system. For consistency
with Shatnawi, we explore the correlation between
the code metrics and the defect counts and   reject code metrics that are poor predictors of defects (i.e.   those  with $p > 0.05$).
Following this, the normalized metric values are ordered in an ascending fashion (this is equivalent to computing a density function, in which the x-axis represents the weight ratio (0-100\%), and the y-axis the metric scale).
Alves et al. then compute their 
thresholds are then derived by choosing the percentage of the overall code that needs to be represented. For instance, Alves et al. suggest the use 90\% quantile of the overall code to derive the threshold for a specific metric. The intuition here is that the worst
code are the outliers beyond 90\% of the normal code measurements.

@@

\begin{figure*}[!b]
~\hrule~
\begin{minipage}{.53\linewidth}
\small
 \begin{tabular}{p{0.95\linewidth}} 
 On the right-hand-side is a tree
 generated by iterative dichomization. 
 This tree can be read like a nested if-then-else statement; e.g.
 \begin{itemize}
     \item Lines 3 and 8 show two branches for lines of code (denoted here as {\em `\$loc}) below 698 and above 698.
     \item Any line with a colon ":" character shows  a  leaf  of  this  nesting.   For  example,  if  some  new  code module is passed down this tree and falls to the line marked in \textcolor{orange}{{\bf orange}},  the colon on that line indicates a prediction that this module has a 100\% chance of being defective. 
     \end{itemize}
Using this tree,
 XTREE looks for a nearby branch that has a lower chance of being defective. Finding the \textcolor{green}{{\bf green}} desired  branch,  XTREE reports a bad smell threshold for that module that is the delta between the \textcolor{orange}{{\bf orange}}  current branch  and   \textcolor{green}{{\bf green}} designed branch. 
 In this case, that threshold relates to:
 \begin{itemize}
     \item Lines of code and comments ({\em lcom}) 
     \item The cohesion between classes ({\em cam}) which measures similarity of parameter lists to assess the relatedness amongst class   methods.
     \end{itemize}\\ 
 \end{tabular}
 \end{minipage}~~~~
 \begin{minipage}{.45\linewidth}
\includegraphics[width=\linewidth]{figs/XTREE_samp.png}
\end{minipage}
~\hrule~
 \caption{A brief tutorial on XTREE.} \label{fig:xtree_samp}
\end{figure*}

\subsubsection{Discussion of the Outlier Approach}
The advantage of the outlier
approach is that it is simple to implement. 
For example, in less than a day we could
code up and apply all the approaches of Shatnami and 
Alves et al.

On the other hand, these outlier approaches have major two disadvantages. 
Firstly, they are {\em verbose}; i.e.  they can
propose thresholds on most of the code metrics.
Such verbosity makes it hard   to critique and prune bad smells
since any bad smell mentioned by a developer has some support
somewhere amongst all those proposed thresholds.

Secondly, the outlier approaches described above are all 
{\em univariant} which means
they do  not understand
that sets of code metrics must
be changed at the same time.
For example:
\begin{itemize} 
\item One way to remove unusually large methods is to refactor that method's
functionality into several smaller ones. 
\item If that is done,
this means that {\em reducing} lines of code in one method
will {\em increase} the coupling of the methods of that class to
either itself or to other classes (depending on whether or not
the new smaller methods are placed in the original class, or in other classes).
\item
Hence, it can be counter-productive to act on bad smell thresholds for {\em both}
large lines of code {\em and} seemingly excessive coupling since
one cannot be reduced without increasing the other.
\end{itemize}

@@

\subsubsection{CD}
Borges and Menzies first proposed the CD centroid delta approach to avoid the drawbacks of univariant reasoning~\cite{me12c}.
Centroid deltas are a {\em multivariant} method which
work  as follows:
\begin{itemize}
    \item Cluster data about code modules;
    \item Find
neighboring clusters $C_+,C_-$ where $C_+$ has more examples of defective
modules than $C_-$;
\item Compute the  delta   in code metrics between the clusters using \mbox{$\Delta = C_+ - C_-$};
\item $\Delta$ are changes needed in defective modules of $C_+$ to
      make them more like the less-defective modules of $C_-$
\end{itemize}
Note that $\Delta$ is a multivariant recommendation
that proposes multiple changes in order to reduce defects.
Since it is computed
from neighboring clusters, it   reasons about
examples with similar distributions.  Hence,   $\Delta$  respects the naturally occurring constraints in the data. For example,
repeating the above example about large methods,   $\Delta$   will not  suggest lowering lines of code
without also increasing a coupling measure.

The drawback with with CD is that it is very {\em verbose}
since
CD   recommended changes to all code
metrics with different values in those two centroids. 
This makes it hard to use CD to   critique and prune away bad smells.
\subsubsection{XTREE}

@@

  Decision trees are built via
  {\em iterative dichotomization} (ID).
  ID   finds  a  split  in  the  values  of  independent  features  that  most  reduces the variability of the dependent feature. The data is then divided on that split
  and ID is called
  recursively on the data in each split. In this way,
  each split is the root of a sub-tree.

  The CART iterative dichotomization algorithm
  measures the variability of continuous attribites
  using the  standard deviation $\sigma$ of the defect counts.
  If a code metric is split into $r$ ranges and each range is of
  size
    $n_r$ is
  associated with a set of defect counts $x$ with standard deviation
  $\sigma_r$ then the best split is the one that minimizes
  $\sum_r\frac{n_r}{n}\sigma_x$ (where $n=\sum_r n_r$).

>>>>

\documentclass{sig-alternate}
%\usepackage{stfloats}
\usepackage{tikz}
  \def\firstcircle{(90:1.75cm) circle (2.5cm)}
  \def\secondcircle{(210:1.75cm) circle (2.5cm)}
  \def\thirdcircle{(330:1.75cm) circle (2.5cm)} 
\usepackage{comment}
\usepackage{cite}
\usepackage{framed,graphicx,xcolor}
\usepackage{multirow}
\usepackage{rotating}
%\usepackage{stfloats}

@@

This paper proposes   XTREE, a tool to   prioritize  bad  smells.
Given an historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have little or no support in that log.

When XTREE's recommendations are compared to 
  those found by previous work (Shatnawi, Alves, and Borges.)
  If code modules are  changed to avoid the bad
  smells found by XTREE,  that changed system
  has significantly fewer expected defects.
  Also, XTREE   endorses
  changes to far fewer code metrics.
  Hence, it is 
  a better tool for ruling out irrelvant bad smells.
  For example, the XTREE results show that many of the bad smells learned 
  by Shatnawi, Alves, and Borges can be ignored.

@@

On the other hand, other papers cast doubts on the value of use bad smalls
as triggers for code change.
As we show in this paper,  
 tools and text books and developers can disagree on what bad smells
are important enough to hunt down and eliminate. Our findings
are consistent with other research suggesting that bad
smells may not be a useful way to guide refactoring~\cite{Mantyla2004,Yamashita2013,Sjoberg2013}.

@@

To address this problem, we propose the XTREE
tool to judge refactorings
proposed by developers.
  XTREE requires:
     \begin{itemize}
         \item An historical log of defects seen previously in the code;
         \item  Bad smells expressed in terms of the 
          metrics   in that log;
         \item Some proposed refactorings justified using those bad smells.
     \end{itemize}
Given that information,  XTREE can determine which
  proposed refactorings are be ignored
since they are not justifiedable  
with respect to the history of this project.

XTREE addresses four problems that  make it hard to  learn useful
bad smells for a particular project.

{\em PROBLEM \#1:} As discussed in the next section, programmer
   cognitive biases result in different
developers declaring that
bad smell X or Y is most important. 
 XTREE offers a way to assess
 competing claims about different refactorings.

{\em PROBLEM \#2:} Not all bad smells can be fixed. 
Consider
a bad smell that is triggering on some very long method,
yet that method must be passed as a single closure
to a third-party GUI toolkit.  This is one
example of a common problem-- when new code must
be combined with code from other developers,
sometimes the constraints imposed by the other code
lead unavoidably to bad smells. In this case,
a refactoring proposed to reduce that bad smell
may be irrelevant since it cannot be applied to this code base.

To address this second
  problem, we let
XTREE {\em reason optimistically} to find the set
$R_0$ of refactorings that might be useful
(for details on how that optimistic reasoning
is implemented, see \tion{rq}).
Since that set was generated optimistically,
the space of actually usable refactorings is 
the possibly smaller subset $R_1 \subseteq R_0$.
If the developer's proposed refactoring $R_2$ is
not in $R_0$, then it cannot be in the smaller
set $R_1$.  As shown
in this paper's experiments,  XTREE's
recommendations typically include only a small subset of
the code metrics. That is, we
can still rule out some proposed refactorings
since many of those proposals fall outside of the space
of refactorings suggested by XTREE.

{\em PROBLEM \#3:}  
Shantnawi~\cite{Shatnawi10} warns most  studies do not offer
{\em both} a metrics and a threshold for their proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
discuss what OO design metrics might defect bad smells (e.g. a excessive
depth of inheritance tree) but they never comment on what thresholds
means that some section of code starts to smell bad.  
In all our  reading of the recent literature,
the only work we can find from other researchers that
offers metrics {\em and} thresholds come
from Shatnami~\cite{Shatnawi10}
and  Alves et al.~\cite{Alves2010} (for details on those
methods, see \tion{methods}).
One potential problem with their approaches is that 
they do not explore the interconnections between sets
of code metrics. This is important if sets of
code metrics have to be changed at the same time.
For example, it is not possible to refactor a
long method it several smaller ones without increasing
the coupling metrics of a class. 
More generally, we say that   bad smell thresholds should be
{\em conjunctions} of metric+thresholds pairs that need to be changed {\em togehter}.
Yet Shatnami~\cite{Shatnawi10}
and Alves et al.   report   {\em disjunctions} of multiple metric/thresholds~\cite{Shatnawi10,Alves2010}
which could be separately and  inappropriately used
to guide refactoring.

To address this third problem, 
XTREE   generates {\em conjunctions} of   thresholds
than should  be considered at the same time.

{\em PROBLEM \#4:} It is hard to judge the  effects of removing bad smells.
Refactored code cannot be assessed just by a rerun of the the test
suite since
code refactorings is meant to
preserve the semantics of the code (i.e. the same tests
should pass before and after the refactoring). 
Also, it many not be practical
 to assessing refactorings by making the
refactorings,  then seeing what happens to that code for the rest
of the project. In a test-driven agile project, as
SCRUM meetings change what stories are implemented next,
the operational profile that lead to defects in the old
code may not be repeated in future executions. 
To resolve this  problem, SE researchers such as 
Cheng et al.~\cite{Cheng10}, O'Keefe et al.~\cite{OKeeffe08,OKeeffe07},
Moghadam~\cite{Moghadam2011} and Mkaouer et al.~\cite{Mkaouer14}
use a pair of systems: 
\begin{enumerate}
    \item One system to propose changes;
\item Another, completely separate, system to assess
how defective is the code before and after some
refactoring.
\end{enumerate}
For the second system,
 Cheng, O'Keefe, Moghadam and  Mkaouer et al. use the QMOOD hierarchical
quality model~\cite{Bansiya02}, constructed in 2002.
Our problem with using tools like QMOOD
is that   quality models learned from other projects,
may perform poorly when applied to new projects~\cite{me12d}.

To solve this fourth problem, this study  eschews
older quality models like QMOOD.
Instead, we use
Random Forests~\cite{Breiman2001} to learn defect predictors
that are specific to the systems   studied here.

 As to XTREE's limitations, the tool 
  can only comment
on refactorings expressed in the   terminology
as the data
  in the historical log of a project.
If developers want to justify their refactorings
via bad smells expressed in other terminology,
then a XTREE-style analysis must   wait till 
data about those new
terms has been collected.

Another  aspect to note about XTREE  is that the tool assumes
Fowler's bad smells
should be assessed only via their contribution to software defects
(evidence: see the section in italics of the quote
that starts this introduction). Some researchers 
 comment that   debating bad smells is  useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
debates about refactorings can   (a)~introducing new comers to the code base or (b)~share knowledge around a team
about interactions within the code. In that view, 
the concept of refactoring can be useful
even if it does not address issues of defects. 
That said, using
defect predictors,
it is possible to operationalize XTREE
while it might be harder to otherwise operationalize the assessment of bad smells.
Hence, 
this paper assumes that {\em better} bad smells and
bettering at finding code changes that  {\em reducing} the chances
that the new code will contain defects.

The rest of this paper is structured as follows. XXX.
discusses the cognitive
biases which make developers disagree on the relevance of those bad smells to
their particular projects.  
 That discussion is our motivation for developing the XTREE  tool
that   carefully assesses the merit of old bad smells to new projects.  XTREE is presented in Section 3 and evalauted in the rest of the paper.

\subsection{Contributions of this Paper }
The XTREE learner has not been reported
previously.
 The experimental rig
 used to comparatively assess
    XTREE is another novel contribution. Using the design
    of that rig,
    other researchers could confirm/improve/refute
    the conclusions of this paper.
    Another contribution here is
    the comparative analysis of XTREE
    versus previously published approaches to learning
    bad smell detectors. that comparion shows
    that XTREE   rarely perfoms
      worse that the other approaches~\cite{Shatnawi10,Alves2010}
      and often is it   much better at proposing
      changes that reduce the expected
      value of the defects. Further, the number
    of code metrics mentioned by XTREE is far
    fewer than other approaches This means that XTREE is a better
    tool for ruling out   refactorings proposed by
    developers.

As to the relation to prior work, we
show here that XTREE is a  superior 
to the CD tool   developed previously with Borges~\cite{me12c}.

@@

Two of the studies in \fig{smells} offers some comments on the relative importance
of the different bad smells the   Yamashita'13 study and the developer study.Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey. Those three were {\em duplicated code, large class}, 
        and {\em long method}. 
 Note that this agreement also means that the
  Yamashita'13 study and the {\em developer survey}   
  believe
        that very few  code smells are   high priority issues
        requiring refactoring.

In summary, just because a developer strongly believes something
about bad smells, does
not necessary mean that that belief is useful for the current project.
Human developers can be clever, but their thinking can also be distorted
by cognitive biases.
Hence, as shown in \fig{smells}, developers and text books and tools 
can disagree on which bad smells are important.
Special tools are needed to assess their beliefs in, for example,
bad smells.

\section{Learning Bad Smell Thresholds}
The SE literature offers  two ways to 
learning bad smell thresholds.
One way relies on 
{\em outlier statistics}~\cite{erni96,bender99}. This way
has been used   by Shatnawi~\cite{Shatnawi10}, Alvenes et al.~\cite{Alves2010}
    and Hermans et al.~\cite{hermans15}.
    Another approach is 
    based on {\em cluster deltas} that we developed
    for   Centroid Deltas~\cite{me12c} and 
    use here for XTREE.

These two approaches are discussed below.
\subsection{  Outliers}\label{sect:p}

@@

\subsubsection{Enri \& Lewerentz}
One of the first methods was proposed Erni and Lewerentz~\cite{erni96}. 
Given classes described with the  code metrics of \fig{ck},
Enri and Lewerentz found the   mean $\mu$ and the standard deviation $\sigma$
of each
code metrics. Their definition of problematic outlier was any code
metric with a measurement greater than $\mu+\sigma$.

Shatnami and Alves et al.~\cite{Shatnawi10,Alves2010} depreciate
using $\mu+\sigma$. since it does not   consider the fault-proneness of classes when the thresholds are computed. Also, the method lacks  empirical validation.

\subsubsection{ Shatnami}
Shatnami~\cite{Shatnawi10}'s preferred alternative to $\mu+\sigma$
is to use the VARL method (Value of Acceptable Risk Level) proposed by Bender~\cite{bender99} in his epidemiology studies.  This approach uses two
thresholds which, following Shatnami's guidance, we set to
$p_0=p_1=0.05$.

VARL encodes the defect count
for each class to 0 (for no defects known) to 1 (for defects known).
Univariate binary logistic regression is applied to learn three coefficients:  
 $\alpha$ is  constat intercept;
    $\beta$ is the coefficient for maximizing log-likelihood;
  and $p_0$  
    measures   how well this   model predicts for   defects.
Any code metric with $p_0>0.05$ is  rejected as being a poor defect predictor. Thresholds are then learned from the surviving metrics $M_c$ using
an equation proposed by Bender:
\begin{equation}
 \mathit{bad\; smell\; if\;} M_c > \frac{1}{\beta }\left( {\log \left( {\frac{{{p_1}}}{{1 - {p_1}}}} \right) - \alpha } \right) 
\end{equation}
\subsubsection{ Alves et al.}
Alves et al.~\cite{Alves2010} propose another approach
to finding thresholds that  utilizes the underlying statistical distribution and scale of the metrics. 
Evey metric value is weighted according to the source lines of code (LOC) of the class. All the weighted metrics are then normalized; i.e. they are divided by the sum of all weights of the same system. For consistency
with Shatnawi, we explore the correlation between
the code metrics and the defect counts and   reject code metrics that are poor predictors of defects (i.e.   those  with $p > 0.05$).
Following this, the normalized metric values are ordered in an ascending fashion (this is equivalent to computing a density function, in which the x-axis represents the weight ratio (0-100\%), and the y-axis the metric scale).
Alves et al. then compute their 
thresholds   by choosing the percentage of the overall code that needs to be represented. For instance, Alves et al. suggest the use 90\% quantile of the overall code to derive the threshold for a specific metric. The intuition here is that the worst
code have outliers beyond 90\% of the normal code measurements.

@@

\begin{figure*}[!t]
~\hrule~
\begin{minipage}{.53\linewidth}
\small
 \begin{tabular}{p{0.95\linewidth}} 
 On the right-hand-side is a tree
 generated by iterative dichomization. 
 This tree can be read like a nested if-then-else statement; e.g.
 \begin{itemize}
     \item Lines 3 and 8 show two branches for lines of code (denoted here as {\em `\$loc}) below 698 and above 698.
     \item Any line with a colon ":" character shows  a  leaf  of  this  nesting.   For  example,  if  some  new  code module is passed down this tree and falls to the line marked in \textcolor{orange}{{\bf orange}},  the colon on that line indicates a prediction that this module has a 100\% chance of being defective. 
     \end{itemize}
Using this tree,
 XTREE looks for a nearby branch that has a lower chance of being defective. Finding the \textcolor{green}{{\bf green}} desired  branch,  XTREE reports a bad smell threshold for that module that is the delta between the \textcolor{orange}{{\bf orange}}  current branch  and   \textcolor{green}{{\bf green}} designed branch. 
 In this case, that threshold relates to:
 \begin{itemize}
     \item Lines of code and comments ({\em lcom}) 
     \item The cohesion between classes ({\em cam}) which measures similarity of parameter lists to assess the relatedness amongst class   methods.
     \end{itemize}\\ 
 \end{tabular}
 \end{minipage}~~~~
 \begin{minipage}{.45\linewidth}
\includegraphics[width=\linewidth]{figs/XTREE_samp.png}
\end{minipage}
~\hrule~
 \caption{A brief tutorial on XTREE.} \label{fig:xtree_samp}
\end{figure*}

\subsubsection{Discussion }\label{sect:disc}
The advantage of the outlier
approach is that it is simple to implement. 
For example, in less than a day we could
code up and apply  the Shatnami and 
Alves et al. methods.

The outlier approaches has   two  major disadvantages. 
Firstly, they are {\em verbose}; i.e.  they can
propose thresholds on most of the code metrics.
Such verbosity makes it hard   to critique and prune bad smells
since any bad smell mentioned by a developer has some support
somewhere amongst all those proposed thresholds.

Secondly, the outlier approaches suffers from
the {\em disjunction} problem discussed in the introduction.
That is, while
 they propose thresholds for many code metrics
individually, they make no comment on what
metrics need to be changed at the same time.

@@

Cluster deltas are a general method
for learning {\em conjunctions} of changes
that need to be applied at the same time. 
This approach works as follows:
\begin{itemize}
    \item Cluster the data. 
    \item Find
neighboring clusters $C_+,C_-$ where $C_+$ has more examples of defective
modules than $C_-$;
\item Compute the  delta   in code metrics between the clusters using \mbox{$\Delta = C_+ - C_-$};
\item $\Delta$ are changes needed in defective modules of $C_+$ to
      make them more like the less-defective modules of $C_-$
\end{itemize}
Note that $\Delta$ is a conjunction of  recommendations.
Since it is computed
from neighboring clusters, it   reasons about
examples with similar distributions.  Hence,   $\Delta$  respects the naturally occurring constraints in the data. For example,
repeating the above example about large methods,   $\Delta$   will not  suggest lowering lines of code
without also increasing a coupling measure

Cluster deltas are used in CD~\cite{me12c} and XTREE.

\subsubsection{CD}
Borges and Menzies first proposed the CD centroid delta approach to
generate {\em conjunctions} of code metrics
that need to be changes at the same time
in order to reduce defects~\cite{me12c}.
CD used the WHERE clustering algorithm developed by the
authors for a prior application~\cite{localvsglobal}.
Each cluster was then replaced by its centroid
and $\Delta$ was calculated directly from the difference
between code metrics values between one centroid
and its nearest neighbor.

As shown below, one drawback with with CD is that it is very {\em verbose}
since
CD   recommended changes to all code
metrics with different values in those two centroids. 
This makes it hard to use CD to   critique and prune away bad smells.

Another drawback with CD is revealed by the comparison experiments
in this paper: CD never worked as well as XTREE in proposing
changes to reduce defects.

\subsubsection{XTREE: Overview}

@@

  Decision trees are built via
  {\em iterative dichotomization} that  finds  a  split  in  the  values  of  independent  features  that  most  reduces the variability of the dependent feature. The data is then divided on that split
  and iterative dichotomization is called
  recursively on the data in each split. In this way,
  each split is the root of a sub-tree.

  The CART iterative dichotomizer~\cite{breiman84}
  measures   variability  
  using the  standard deviation $\sigma$ of the defect counts.
  If a code metric is split into $r$ ranges and each range is of
  size
    $n_r$ is
  associated with a set of defect counts $x$ with standard deviation
  $\sigma_r$ then the best split is the one that minimizes
  $\sum_r\frac{n_r}{n}\sigma_x$ (where $n=\sum_r n_r$).

  While not usually labelled as such, CART can
  be viewed as a clustering algorithm that groups together code modules
  with similar defect counts and some shared ranges of code metrics.
  XTREE is a post-processor to CART that reports the delta between 
  nearby groups with different defect counts. \fig{xtree_samp} offers
  a small example of that reflection. In this example, some project has a table of data describing its  classes in the form of \fig{ck}. After code inspections and running test cases or operational
      tests, each such class is augmented with a defect count.
      CART takes that table of data and, 
        generates the tree of \fig{xtree_samp}.

     Once the tree is built, some test case arrives and is passed
      down the tree to a leaf note (see the \textcolor{orange}{{\bf orange}} line in \fig{xtree_samp}.
       XTREE then looks for a nearby leaf node with a lower defect
      count (see the \textcolor{green}{{\bf green}} line in \fig{xtree_samp}). For that test case, XTREE proposes a bad smell
      thresholds that is  the difference between 
      \textcolor{green}{{\bf green}} and \textcolor{orange}{{\bf orange}}.

\subsubsection{XTREE:   Details}

Using the training data,  XTREE uses CART to
  divide  $N$ code modules  into  clusters of
size $\alpha=\sqrt{N}$.
